The code
===

This `shiny` example illustrates the idea of interactive data exploration. For background on the public transportation dataset used, see this [github repository](https://github.com/tcrug/public-transpo). User inputs reactively set the variables to be used by `ggplot()` for the x and y axes as well as, optionally, color and size. In addition, the user is able to apply transformations (`log()` and `sqrt()`) to the axis and size variables.

**ui.R**

```{r}
library(shiny)
library(markdown)

shinyUI(pageWithSidebar(
  # title
  headerPanel("Transportation exploration"),
  
  sidebarPanel(
    
    selectInput(inputId = "x",
                label = "Select x-axis variable", 
                list("Population" = "population", "Service area" = "service_area_sq_mi",
                     "VOMS" = "voms", "Annual miles" = "annual_vehicle_miles",
                     "Vechicle hours" = "annual_vehicle_hours", "Unlinked trips" = "unlinked_passngr_trips",
                     "Passenger miles" = "passenger_miles", "Total BTUs" = "btus_total",
                     "BTUs/passenger-mile" = "btus_pmile", "Ave. occupancy" = "average_occupancy")),
    
    selectInput(inputId = "y",
                label = "Select y-axis variable", 
                list("Population" = "population", "Service area" = "service_area_sq_mi",
                     "VOMS" = "voms", "Annual miles" = "annual_vehicle_miles",
                     "Vechicle hours" = "annual_vehicle_hours", "Unlinked trips" = "unlinked_passngr_trips",
                     "Passenger miles" = "passenger_miles", "Total BTUs" = "btus_total",
                     "BTUs/passenger-mile" = "btus_pmile", "Ave. occupancy" = "average_occupancy")),
    
    selectInput(inputId = "color",
                label = "Select color variable", 
                list("None" = "none", "City" = "city", "State" = "state", "Org" = "org_type",
                "Mode" = "mode", "Service" = "service", "Energy type" = "primary_energy")),
    
    selectInput(inputId = "size",
                label = "Select size variable", 
                list("None" = "none", "Population" = "population", "Service area" = "service_area_sq_mi",
                     "VOMS" = "voms", "Annual miles" = "annual_vehicle_miles",
                     "Vechicle hours" = "annual_vehicle_hours", "Unlinked trips" = "unlinked_passngr_trips",
                     "Passenger miles" = "passenger_miles", "Total BTUs" = "btus_total",
                     "BTUs/passenger-mile" = "btus_pmile", "Ave. occupancy" = "average_occupancy")),
    
    selectInput(inputId = "x_trans", label = "x variable transformation",
                list("None" = "identity", "log" = "log", "sqrt" = "sqrt")),
    
    selectInput(inputId = "y_trans", label = "y variable transformation",
                list("None" = "identity", "log" = "log", "sqrt" = "sqrt")),
    
    selectInput(inputId = "size_trans", label = "Size variable transformation",
                list("None" = "identity", "log" = "log", "sqrt" = "sqrt"))
  ),
  
  mainPanel(
    tabsetPanel(
      tabPanel("Plot",
               plotOutput("plot", width = "100%", height = "600px")),
      tabPanel("Code",
               includeMarkdown("./code.Rmd"))
  )
)))
```

**server.R**
  
```{r}
library(shiny)
library(ggplot2)

data <- read.csv("public-transpo.csv")

# need a color to use when none is chosen
data$none <- rep("", nrow(data))

shinyServer(function(input, output) {
    
  output$plot <- renderPlot({
    
    # initialize ggplot; aes_string is necessary since I don't know how to 
    # pass objects from the input object. Instead, intpu$var returns a string
    # so we use the handy aes_string() method instead of regular 'ol aes()
    p <- ggplot(data, aes_string(x = input$x, y = input$y))
    
    # there were issues using range(min, max) in scale_size below, so I opted to
    # just run a check to see if no size variable had been selected.
    # If not, we just use a constant value.
    if(input$size == "none") {
      p <- p + geom_point(aes_string(colour = input$color), size = 4)
    }
    
    # if size is not "none", add the aesthetic
    else {
      p <- p + geom_point(aes_string(size = input$size,
                                     colour = paste0("as.factor(", input$color, ")")))
    }
    
    # ggplot takes trings for transformations, so we can handily just pass
    # the appropriate name right to the scale!
    p <- p + scale_x_continuous(trans = input$x_trans)
    p <- p + scale_y_continuous(trans = input$y_trans)
    
    # the range() option is nice so we can make sure the min and max bubble
    # sizes are reasonable. We can also apply a transformation here.
    p <- p + scale_size_continuous(input$size,
                                   trans = input$size_trans, range = c(5, 25))
    
    # just unifying the color name so we don't get as.factor()
    p <- p + scale_colour_discrete(input$color)
    
    # some theme options to pretty things up and make it more readable
    p <- p + theme_bw() + theme(text = element_text(size = 18))
    
    print(p)
    
  })
})
```